
#Making ranked plot

load relevant libraries and read consensusfasta file

```{r}
if (TRUE){
    setwd("~/Dropbox/MarionKristofBachelerProject/CodePleuni/")
	library(ape)
	library(seqinr)
    #install.packages("pegas")
    library(pegas)
	#read the file with the resistance mutations
	source("RResistanceMutations.r")
	#read the fasta file 
	consensusfasta<-read.dna("HIV1_CON_2004_POL_DNA.fasta", format = "fasta",as.character=TRUE)	
	#where is the start of POL? 
	polstart=regexpr("cctca",paste(consensusfasta[which(row.names(consensusfasta)=="CONSENSUS_B"),],collapse=""))[1]
	consensusB<-consensusfasta[which(row.names(consensusfasta)=="CONSENSUS_B"), polstart:(polstart+983)]
	consofcons<-consensusfasta[which(row.names(consensusfasta)=="CON_OF_CONS"), polstart:(polstart+983)]
}
```

#Read the stored frequencies in stead
```{r}
if (TRUE){#read the stored data
read.table("../freqPatSiteInclDay0.csv",sep=",",header=TRUE,row.names=1)->freqPatSite
read.table("../freqPatTsInclDay0.csv",sep=",",header=TRUE,row.names=1)->freqPatTs
}
```
```{r}
transition<-function(nuc){
    if (nuc=="a") return("g")
    if (nuc=="g") return("a")
    if (nuc=="c") return("t")
    if (nuc=="t") return("c")
}
```
#for each mutation, determine whether it is synonymous, non-synonymous or creates a stop

```{r}
typeofsitefunction<-function(WTcodon, mutantcodon){
    WTAA<-translate(WTcodon)
    MUTAA<-translate(mutantcodon)
    if (WTAA == MUTAA) return ("syn")
    else if (MUTAA == "*") return ("stop")
    else return ("nonsyn")
}
```    

```{r}
TypeOfSite<-c()
for (codon in 1:(984/3)){#for each codon in the sequence
    positions <- c(codon*3-2,codon*3-1, codon*3)
    WTcodon <- consensusB[positions]
    mutant1codon <- c(transition(WTcodon[1]), WTcodon[2:3])
    mutant2codon <- c(WTcodon[1],transition(WTcodon[2]), WTcodon[3])
    mutant3codon <- c(WTcodon[1:2], transition(WTcodon[3]))
    TypeOfSite<-c(TypeOfSite,typeofsitefunction(WTcodon,mutant1codon))
    TypeOfSite<-c(TypeOfSite,typeofsitefunction(WTcodon,mutant2codon))
    TypeOfSite<-c(TypeOfSite,typeofsitefunction(WTcodon,mutant3codon))
}
```    

```{r}
#make sure that resistance sites in RT have a diff type of site
TypeOfSite[sort(c((RTImuts$pos*3)-2,(RTImuts$pos*3)-1,(RTImuts$pos*3)))+297]<-"res"
```

```{r}
#Calculate mean freq for each site 
attach(freqPatSite); colMeans<-apply(freqPatSite, 2 , mean, na.rm=TRUE)
attach(freqPatTs); colMeansTs<-apply(freqPatTs, 2 , mean, na.rm=TRUE)

#Get conf intervals by bootsstrapping
btmeansTs<-data.frame(row.names=names(freqPatTs)[1:984]) #each row is a site, each column is a bootstrapped mean
numbootstraps=100 # small number to make it faster
for (j in 1:984){# start with the first site
for (k in 1:numbootstraps){ # first iteration
  btmeansTs[j,k]=mean(sample(freqPatTs[,j],length(freqPatTs[,j]),replace = TRUE),na.rm=TRUE)
}}
btmeansSorted<-t(apply(btmeansTs,1,sort))
lowerConf<-btmeansSorted[,floor(0.025*numbootstraps)]
upperConf<-btmeansSorted[,floor((1-0.025)*numbootstraps)]

x<-data.frame(num=1:984,colMeansTs,TypeOfSite,lowerConf,upperConf)
newdata <- x[order(colMeansTs),] 
newdata$color<-""
for (i in 1:984){
    if (newdata$TypeOfSite[i]=="syn") newdata$color[i] = "green"
    if (newdata$TypeOfSite[i]=="nonsyn") newdata$color[i] = "red"
    if (newdata$TypeOfSite[i]=="stop") newdata$color[i] = "black"
    if (newdata$TypeOfSite[i]=="res") newdata$color[i] = "purple"
}
```

```{r}
#pdf("Protease.pdf",width = 13, height = 10)
PROdata<-newdata[newdata$num<298,]
#remove resistance mutations
PROdata<-PROdata[PROdata$TypeOfSite!="res",]
plot(PROdata$colMeansT+0.001,log="y",ylim=c(0.001,0.15),cex=2, pch = "|",col=PROdata$color)
#arrows(1:length(PROdata$num), PROdata$lowerConf +0.001, 1:length(PROdata$num), PROdata$upperConf +0.001, 
#       angle=90, length=0.01, code=3, col=PROdata$color,lwd=0.1)
#dev.off()
```

```{r}
#pdf("RT.pdf",width = 13, height = 10)
RTdata<-newdata[newdata$num>=298,]
RTdata<-RTdata[RTdata$TypeOfSite!="res",]
plot(RTdata$colMeansT+0.001,log="y",ylim=c(0.001,0.15),cex=1, pch = "|",col=RTdata$color)
#arrows(1:length(RTdata$num), RTdata$lowerConf +0.001, 1:length(RTdata$num), RTdata$upperConf +0.001, 
#       angle=90, length=0.01, code=3, col=RTdata$color,lwd=0.1)
#dev.off()

#For protease, which are the not-lethal stop codons? 
```{r}
#head(PROdata)
PROdata[PROdata$TypeOfSite=="stop"& PROdata$colMeansTs>0,]
#site 4 and site 125
#Which codons are they part of?
plot(PROdata$colMeansT+0.001,log="y",ylim=c(0.001,0.2),cex=2, pch = "|",col=PROdata$color)

sites<-PROdata$num[PROdata$TypeOfSite=="stop"& PROdata$colMeansTs>0]
codons<-ceiling(sites/3)

for (codon in codons){
    positions <- c(codon*3-2,codon*3-1, codon*3)
    rows<-which(PROdata$num %in% sites)
    points(rows,PROdata$colMeansT[rows]+0.001,col=PROdata$color[rows],cex=2,pch=16)
#    print(positions)
#    WTcodon <- consensusB[positions]
#    print (WTcodon)
#    print(translate(WTcodon))
#    print(PROdataz#print frequencies
}

#points(PROdata$colMeansT+0.001)

```

```{r}
#for ease of reading 
freqPatTs<-round(freqPatTs,2)
```

#I need to look at the original fasta files and look at the frequency of codons. 
```{r}
list.files(path="../FASTAfiles/")->listfastafiles
for (co in codons){#for each site in the sequence
    ListAA<-c(); ListCodon<-c()
    for (i in 1:length(listfastafiles)){ #for each fastafile 
        filename=paste("/Users/pleuni/Dropbox/DevanShahProject/FASTAfiles/",substr(listfastafiles[i],1,6),".fasta",sep="")
	    #print(filename)
	    patfasta<-read.dna(filename, format = "fasta",as.character=TRUE) #read the file 
		if (length(patfasta[,1])>1) {
        WTcodon =	consensusB[c(co*3-2,co*3-1,co*3)]; #what is WT at site j?
        WTAA = translate(WTcodon)
        Patcodon = patfasta[,c(co*3-2,co*3-1,co*3)]
        #print(Patcodon)
        PatAA = apply(Patcodon, 1, function(x) translate(x))
        ListAA<-c(ListAA, PatAA)
        Patcodon<-apply(Patcodon, 1, function(x) paste(x,collapse=""))
        ListCodon<-c(ListCodon,Patcodon)
        }
    }
    print(paste("Codon: ", co))
    print(table(ListAA))
    print(table(ListCodon))
}

```
# how do they get these high frequencies. Is it due to a few patients? 
#So for the first stop codon, the stop codon actually occurs, but in very small numbers (count: 3)
#For the second stop codon, the polymorphism that would create a stop codon is always paired with a neighboring mutation that makes it a non-syn mutation. 

```{r}
PROdataNonSyn<-PROdata[PROdata$TypeOfSite=="nonsyn",]
#Make plot of Protease 
plot(PROdataNonSyn$num[order(PROdataNonSyn$num)],PROdataNonSyn$colMeansTs[order(PROdataNonSyn$num)]+0.001,col=PROdataNonSyn$color[order(PROdataNonSyn$num)],pch=16,log="y",type="b")
#Add moving mean
for (i in 20:280){
    range = (i-20):(i+20)
#    points(i, mean(PROdataNonSyn$colMeansTs[PROdataNonSyn$num%in%range]), col="blue",pch = 12, cex=3)
    points(i, median(PROdataNonSyn$colMeansTs[PROdataNonSyn$num%in%range]), col="green",pch = 11, cex=3)
}


RTdataNonSyn<-RTdata[RTdata$TypeOfSite=="nonsyn",]
plot(RTdataNonSyn$num[order(RTdataNonSyn$num)],RTdataNonSyn$colMeansTs[order(RTdataNonSyn$num)]+0.001,col=RTdataNonSyn$color[order(RTdataNonSyn$num)],pch=16,log="y",type="b")

#Add moving mean
for (i in 370:900){
    range = (i-20):(i+20)
    points(i, mean(RTdataNonSyn$colMeansTs[RTdataNonSyn$num%in%range]), col="blue",pch = 12, cex=3)
    points(i, median(RTdataNonSyn$colMeansTs[RTdataNonSyn$num%in%range]), col="green",pch = 11, cex=3)
}

i=370
    range = (i-20):(i+20)


```




#For protease, which are the most deleterious non-syn sites? 


#For protease, which are the most deleterious syn sites? 




