---
title: "SynSitesRandomization"
author: "Pleuni"
date: "November 12, 2015"
output: html_document
---


```{r}

source('./baseRscript.R')
```

*Read the stored frequencies rather than calculating frequencies again*
```{r}
if (TRUE){#read the stored data
read.table("../Output/freqPatTsInclDay0-threshold.csv",sep=",",header=TRUE,row.names=1)->freqPatTs
}

colMeansTs<-apply(freqPatTs, 2 , mean, na.rm=TRUE)
OverviewDF<-data.frame(num=1:984,colMeansTs, TypeOfSite)
OverviewDFSyn<-OverviewDF[OverviewDF$TypeOfSite=="syn",]

hist(OverviewDFSyn$colMeansTs,breaks=50)
var(OverviewDFSyn$colMeansTs)

# KT tryout
#removing the % highest proportion from the calculations
#tail(OverviewDFSyn[order(OverviewDFSyn$colMeansTs),],n=-5L)
perc<-0.10
dataSelection<-function(type,percent){
OverviewDFType<-OverviewDF[OverviewDF$TypeOfSite==type,]
perc<-percent/100
npatout<-ceiling(nrow(OverviewDFType)*perc)
OverviewDFTypeOrdered<-OverviewDFType[order(OverviewDFType$colMeansTs),]
OverviewDFTypeOrderedTail<-OverviewDFTypeOrdered[1:(nrow(OverviewDFTypeOrdered)-npatout),]
return(OverviewDFTypeOrderedTail)
}

```

* freqPatTs: data frame of patients (rows) and positions (columns)

* colMeansTs: col means of freqPatTs

* OverviewDF: combination of colmeansts and type of site

+ OverviewDFSyn: all syn positions 

+ OverviewDFSynTail: all syn positions without the x percent highest frequency values 





#OK, now I am going to do the randomization and recalculate the variance 1000 times. Only for the syn sites
=======

### Synonymous positions
OK, now I am going to do the randomization and recalculate the variance 1000 times. Only for the syn sites
```{r}

#original version: no restrictions on outliers
#freqPatTsSyn<-freqPatTs[,OverviewDF$TypeOfSite=="syn"]
#consensusBSyn<-consensusB[OverviewDF$TypeOfSite=="syn"]

#adapted to automate type and percent 
for (w in c(0,2,5,10,15)){
percent<-w  #KT
type<-'syn' #KT
freqPatTsSyn<-freqPatTs[,dataSelection(type,percent)$num] #KT
consensusBSyn<-consensusB[dataSelection(type,percent)$num] #KT

#
Asites<-which(consensusBSyn=="a")
Csites<-which(consensusBSyn=="c")
Gsites<-which(consensusBSyn=="g")
Tsites<-which(consensusBSyn=="t")


ListOfVars<-c()

for (j in 1:1000){
  #  if (j%%10 == 0) print(j)
freqPatTsRandomSyn<-freqPatTsSyn

for (patrow in 1:length(freqPatTs[,1])){
freqPatTsRandomSyn[patrow,sample(Asites)]<-freqPatTsSyn[patrow,Asites]
freqPatTsRandomSyn[patrow,sample(Csites)]<-freqPatTsSyn[patrow,Csites]
freqPatTsRandomSyn[patrow,sample(Gsites)]<-freqPatTsSyn[patrow,Gsites]
freqPatTsRandomSyn[patrow,sample(Tsites)]<-freqPatTsSyn[patrow,Tsites]
}

ListOfVars<-c(ListOfVars, var(apply(freqPatTsRandomSyn, 2 , mean, na.rm=TRUE))) #vector with n=runs means 
assign(paste("ListOfVars",percent,sep="_"),ListOfVars)#KT
}}
```




Show the distribution of variances form the randomized data and compare with the value of the ofiginal dataset. It looks very different!
```{r}
## settings for version 2
#hist(ListOfVars,breaks=20,xlim=c(8*10^-6,2*10^-5), main="Histogram of variances vs observed variance")
## new settings with version 2 adapted with a threshold
min(ListOfVars);max(ListOfVars)

pdf("../Output/Distribution_of_Variances_top0percentout.pdf",width = 13, height = 10)
hist(ListOfVars,breaks=20,xlim=c(.9*10^-5,10*10^-5), main="Histogram of variances vs observed variance (0%)")
abline(v=var(apply(freqPatTsSyn, 2 , mean, na.rm=TRUE)),col=2,lwd=2)
dev.off()
```



Show the distribution of the means for one randome dataset and the original data. One can see that the randomized data have much fewer 0s and fewer high values. 
```{r}
colMeansRandomSyn<-apply(freqPatTsRandomSyn, 2 , mean, na.rm=TRUE)
colMeansSyn<-apply(freqPatTsSyn, 2 , mean, na.rm=TRUE)
pdf("../Output/Distribution_of_Means_top0percentout.pdf",width = 13, height = 10)
par(mfrow=c(1,1))
hist(colMeansSyn,breaks=seq(0,0.2,by=0.001),xlim=c(0,0.04),
     ylim=c(0,40),
     col=rgb(1,0,0,0.5),main="Histogram of mean frequencies 0%")
hist(colMeansRandomSyn,breaks=seq(0,0.1,by=0.001),xlim=c(0,0.04),ylim=c(0,40),add=T,col=rgb(0,0,1,0.5))
legend("topright", c('Randomized','Observed'), lwd=4, col=c(rgb(0,0,1,0.5), rgb(1,0,0,0.5)))
dev.off()
```


#Do the same analysis for the non synonymous sites
```{r}
#tail version: remove the x% upper highest syn positions (=outliers)
freqPatTsSyn<-freqPatTs[,OverviewDFSynTail$num]
consensusBSyn<-consensusB[OverviewDFSynTail$num]

# original version 
freqPatTsNonSyn<-freqPatTs[,OverviewDF$TypeOfSite=="nonsyn"]
consensusBNonSyn<-consensusB[OverviewDF$TypeOfSite=="nonsyn"]
#
AsitesNonSyn<-which(consensusBNonSyn=="a")
CsitesNonSyn<-which(consensusBNonSyn=="c")
GsitesNonSyn<-which(consensusBNonSyn=="g")
TsitesNonSyn<-which(consensusBNonSyn=="t")

ListOfVarsNonSyn<-c()
ListOfNumSites0Freq<-c()
#switch A sites

for (j in 1:100){
  #  if (j %% 10 == 0 ) print (j)
freqPatTsRandomNonSyn<-freqPatTsNonSyn

for (patrow in 1:length(freqPatTs[,1])){
freqPatTsRandomNonSyn[patrow,sample(AsitesNonSyn)]<-freqPatTsNonSyn[patrow,AsitesNonSyn]
freqPatTsRandomNonSyn[patrow,sample(CsitesNonSyn)]<-freqPatTsNonSyn[patrow,CsitesNonSyn]
freqPatTsRandomNonSyn[patrow,sample(GsitesNonSyn)]<-freqPatTsNonSyn[patrow,GsitesNonSyn]
freqPatTsRandomNonSyn[patrow,sample(TsitesNonSyn)]<-freqPatTsNonSyn[patrow,TsitesNonSyn]
}

ListOfNumSites0Freq<-c(ListOfNumSites0Freq,length(which(apply(freqPatTsRandomNonSyn, 2 , mean, na.rm=TRUE)==0)))
ListOfVarsNonSyn<-c(ListOfVarsNonSyn, var(apply(freqPatTsRandomNonSyn, 2 , mean, na.rm=TRUE)))
}
```


Show the distribution of variances form the randomized data and compare with the value of the ofiginal dataset. It looks very different!
```{r}
#hist(ListOfVars,breaks=20,xlim=c(8*10^-6,2*10^-5), main="Histogram of variances vs observed variance")
## new settings with version 2 adapted with a threshold
min(ListOfVarsNonSyn);max(ListOfVarsNonSyn)
hist(ListOfVarsNonSyn,breaks=20,xlim=c(1*10^-6,8*10^-5), main="Histogram of variances vs observed variance")
abline(v=var(apply(freqPatTsNonSyn, 2 , mean, na.rm=TRUE)),col=2,lwd=2)

```
Show the distribution of the means for one randome dataset and the original data. One can see that the randomized data have much fewer 0s and fewer high values. 
```{r}
colMeansRandomNonSyn<-apply(freqPatTsRandomNonSyn, 2 , mean, na.rm=TRUE)
colMeansNonSyn<-apply(freqPatTsNonSyn, 2 , mean, na.rm=TRUE)
par(mfrow=c(1,1))
hist(apply(freqPatTsNonSyn, 2 , mean, na.rm=TRUE),breaks=seq(0,0.2,by=0.001),xlim=c(0,0.04),
     #ylim=c(0,40),
     col=rgb(1,0,0,0.5),main="Histogram of randomized (blue) vs observed (red) mean frequencies")
hist(colMeansRandomNonSyn,breaks=seq(0,0.1,by=0.001),xlim=c(0,0.04),ylim=c(0,40),add=T,col=rgb(0,0,1,0.5))
legend("topright", c('Randomized','Observed'), lwd=4, col=c(rgb(0,0,1,0.5), rgb(1,0,0,0.5)))

```



Just to be sure, I want to check that the non-syn sites have lower frequencies than the syn sites and also the stop sites have lower frequencies than the non-syn sites. 
Not entirely sure what is the best way to do this. 

```{r}
ks.test(colMeansSyn,colMeansNonSyn,alternative="less")

freqPatTsStop<-freqPatTs[,OverviewDF$TypeOfSite=="stop"]
colMeansStop<-apply(freqPatTsStop, 2 , mean, na.rm=TRUE)

ks.test(colMeansNonSyn,colMeansStop,alternative="less")

```

Find out how many non syn sites have 0 frequency and how many we'd expect to have 0 freq. 

```{r}
length(which(colMeansNonSyn==0))

length(which(colMeansRandomNonSyn==0))

hist(ListOfNumSites0Freq,xlim=c(0,30))
abline(v=length(which(colMeansNonSyn==0)),lwd=2,col=2)
```

Where are the non-syn sites with 0 frequency? 

OK, let's redo this with the new data!

```{r}
plot(which(OverviewDF$TypeOfSite=="nonsyn")[which(colMeansNonSyn==0)],rep(1,length(which(colMeansNonSyn==0))),xaxt="n",xlab = "Location",ylab="",ylim=c(0.8,1.2),pch=23)
points(which(OverviewDF$TypeOfSite=="nonsyn")[which(colMeansNonSyn==0)],rep(1,length(which(colMeansNonSyn==0))),pch="|",cex=1)
axis(1, at=c(0,150,297,650,1000), labels=c("","PRO","","RT",""),tck=0)
abline(v=297,col=3,lwd=2)
abline(v=seq(-1,297,by=30),col=3,lwd=1,lty=2)
abline(v=seq(296,1000,by=30),col=2,lwd=1,lty=2)
axis(1, at=seq(-1,297,by=30), labels=seq(0,99,by=10),cex.axis=.5,tck=0,line=-1,lty=0)
axis(1, at=seq(296,1000,by=30), labels=seq(0,234,by=10),cex.axis=.5,tck=0,line=-1,lty=0)
```



=======
par(mfrow=c(1,1))
hist(apply(freqPatTsNonSyn, 2 , mean, na.rm=TRUE),breaks=seq(0,0.2,by=0.001),xlim=c(0,0.04),
     #ylim=c(0,40),
     col=rgb(1,0,0,0.5),main="Histogram of mean frequencies")
hist(colMeansRandomNonSyn,breaks=seq(0,0.1,by=0.001),xlim=c(0,0.04),ylim=c(0,40),add=T,col=rgb(0,0,1,0.5))
legend("topright", c('Randomized','Observed'), lwd=4, col=c(rgb(0,0,1,0.5), rgb(1,0,0,0.5)))
```

#TRY OUT
# from the fasta file consensus
paste(consensusB,collapse="")  

rna<-('.(((((.((.(((((((....)))(((((((((........))).))))))...(((((((..((((((..((((...((((((.((..(((.(((((........(((((((.((((((...(((...))).........((.(((((((((((((((((....)))))))(((((......)))))(((((....)))))..............((((((((.((((((((((...............((((.((((((((((((......))))))..............)))))).)))))))))))))).)))))))).............(((....)))((((((...............))))))..........................(((........)))....(((((......(((...(((((...)))))...))).....)))))......))))))).))).))............((...((((((..........))))))....)).(((..(((.....)))..))).....)))))).)))))))....((....))..(((((...(((((.............(((......(((((((.(((((..(((((((((.((((......(((.....))).......)))).))).))))))..))))).)))))))......)))))).)).)))))..)))))..)))..)).))))))....((((((....((....)).....)))))).................)))))))))))))))))........((((..((((((((((((((......)))))))))...))))).))))(((.((((....)))).)))..............)))))).)))))........((((..((((((......)).)))).))))....................(((......)))')
a<-strsplit(rna,split="")

AA<-data.frame(num=1:984,colMeansTs, TypeOfSite,a[[1]])
names(AA)[4]<-'struc'
AA$struc2<-ifelse(AA$struc=='.','red','blue')
plot(AA$colMeansTs[order(AA$colMeansTs)]+0.00001,pch='|',log='y',col=AA$struc2[order(AA$colMeansTs)])
AAsyn<-subset(AA,AA$TypeOfSite=='syn')
plot(AAsyn$colMeansTs[order(AAsyn$colMeansTs)]+0.00001,pch='|',log='y',col=AAsyn$struc2[order(AA$colMeansTs)])
