#new addition
------
variables of interest
throwawaypatient_1sequence
throwawaypatient_day0seqvar
throwawaypatient_noOflatersequenceslow
numberOfMutations
-----


Calculate the distribution of mutations per codon since in the newest version (V2 and V3), we only allow 1 change per codon, in order to be able associate frequency to outcome. 

For now, remove day0 sequences from the analysis, since they will bias. 


#Load libraries and necessary files from the baseRscript.Rmd
source('baseRscript.R')

#Read the correct fastafiles. 

```{r}
list.files(path="../Data/BachelerFiles/FASTAfiles/")->listfastafiles
lengthlatersequenceslist<-c();lengthallsequenceslist<-c()
```

*make dataframe with frequencies for all non-muts for all patients for all sites* 

```{r}
throwawaypatient_1sequence<-0
throwawaypatient_day0seqvar<-0
throwawaypatient_noOflatersequenceslow<-0
numberOfMutations<-c()
```


```{r}
ListPatientsWithoutData<-c()
for (i in 1:length(listfastafiles)){ #for each fastafile 
filename=paste("../Data/BachelerFiles/FASTAfiles/",substr(listfastafiles[i],1,6),".fasta",sep="")   
	print(filename)
	patfasta<-read.dna(filename, format = "fasta",as.character=TRUE) #read the file 
	
	#which seqs are from first day of sampling
	days=sort(unique(as.numeric(substr(row.names(patfasta),5,7))))
	day0sequences<-which(as.numeric(substr(row.names(patfasta),5,7))==days[1])
	latersequences<-which(as.numeric(substr(row.names(patfasta),5,7))>days[1])
	allsequences=c(day0sequences,latersequences)

	
	if(nrow(patfasta)==1){
  throwawaypatient_1sequence<-throwawaypatient_1sequence+1
	}
	else if(length(latersequences)<2){
  throwawaypatient_noOflatersequenceslow<-throwawaypatient_noOflatersequenceslow+1
	}

	else{
    	for (h in 1:328){  # for each codon 
      codonposition<-((h-1)*3+1):(h*3)
	  
      
      if(FALSE %in% names(table(consensusB[codonposition]==t(patfasta[day0sequences,codonposition])))){
	      throwawaypatient_day0seqvar<-throwawaypatient_day0seqvar+1
	      }
	  else {   #if first sequence -patient codon is equal to consensus codon, then continue  #this is ok
	        numberOfMutations<-c(numberOfMutations,apply(consensusB[codonposition]==(apply(patfasta[latersequences,codonposition], 1, paste)),2,sum))  ## only later sequences.
	  }}}}

	                             
	                              #for (cpos in codonposition){
	        	#	  if (consensusB[cpos]=="c"){freqPatTs_verion3[i,cpos]=length(which(patfasta[good2sequences,cpos]=="t"))/length(good2sequences)}  
		         #   if (consensusB[cpos]=="t"){freqPatTs_verion3[i,cpos]=length(which(patfasta[good2sequences,cpos]=="c"))/length(good2sequences)}
		          #  if (consensusB[cpos]=="a"){freqPatTs_verion3[i,cpos]=length(which(patfasta[good2sequences,cpos]=="g"))/length(good2sequences)}
		           # if (consensusB[cpos]=="g"){freqPatTs_verion3[i,cpos]=length(which(patfasta[good2sequences,cpos]=="a"))/length(good2sequences)}
	     
```


write.csv(numberOfMutations,file='../Output/numberofmutations.csv')


> (table(numberOfMutations))
numberOfMutations
     0      1      2      3 
     7    389  13921 638383 
> prop.table(table(numberOfMutations))*100
numberOfMutations
           0            1            2            3 
 0.001072468  0.059598590  2.132832848 97.806496093 






Determine for each patients, the number of day0 sequences and day1+ sequences, plot them for all 171 patients
```{r}

number_of_day0sequences<-c()
number_of_dayplus1sequences<-c()


ListPatientsWithoutData<-c()
for (i in 1:length(listfastafiles)){ #for each fastafile 
filename=paste("../Data/BachelerFiles/FASTAfiles/",substr(listfastafiles[i],1,6),".fasta",sep="")   
	print(filename)
	patfasta<-read.dna(filename, format = "fasta",as.character=TRUE) #read the file 
	
	#which seqs are from first day of sampling
	days=sort(unique(as.numeric(substr(row.names(patfasta),5,7))))
	day0sequences<-which(as.numeric(substr(row.names(patfasta),5,7))==days[1])
	latersequences<-which(as.numeric(substr(row.names(patfasta),5,7))>days[1])
	allsequences=c(day0sequences,latersequences)

	number_of_day0sequences<-c(number_of_day0sequences,length(day0sequences))	
	number_of_dayplus1sequences<-c(number_of_dayplus1sequences,length(latersequences))
}


percentday0<-number_of_day0sequences/(number_of_dayplus1sequences+number_of_day0sequences)
percentday1<-number_of_dayplus1sequences/(number_of_dayplus1sequences+number_of_day0sequences)


barplot(t(cbind(percentday0,percentday1)),col=c('green','red'))
barplot(t(cbind(number_of_day0sequences,number_of_dayplus1sequences)),col=c('green','red'))
table(number_of_dayplus1sequences>10) #setting a threshold to define a subset?
```








